The effect of restricted substrate diffusion on microbial community function and structure

### 2019-03-25 SI_code.Rmd
MEND restructured

        #equations
        #C uptake
        F1=1/CUE*(Vmax+mr)*Cmic*C/(Km+C)
        #growth respiration
        F4=(1/CUE-1)*Vmax*Cmic*C/(Km+C)
        #maintnance respiration
        F5=(1/CUE-1)*mr*Cmic*C/(Km+C)
        #microbial mortality
        F8=mr*Cmic

        #states
        dCmic<-F1-(F4+F5)-F8
        dC<--F1+F8

        return(list(c(dCmic, dC), r=F4+F5, Protc=0.548*Cmic))

@@

@@

@@

@@

@@

@@

#results of the previous iteration (not run)
# mend_r_results_par<-rbind(mend_r1[[1]]$pars,
#                       mend_r2[[1]]$pars, mend_r2[[2]]$pars, mend_r2[[3]]$pars,
#                       mend_r3[[1]]$pars, mend_r3[[2]]$pars,
#                       mend_r4[[1]]$pars, mend_r4[[2]]$pars, mend_r4[[3]]$pars,
#                       mend_r4[[4]]$pars, mend_r4[[5]]$pars, mend_r4[[6]]$pars)
# mend_r_results_good<-rbind(mend_r1$goodness[2,c(4:7)],
#                      mend_r2$goodness[2,c(4:7)],
#                      mend_r3$goodness[2,c(4:7)],
#                      mend_r4$goodness[2,c(4:7)])
# write.csv(mend_r_results_par, file = c("../mend_r_results_par.csv"), row.names = F)
# write.csv(mend_r_results_good, file = c("../mend_r_results_good.csv"), row.names = F)

@@

@@

        #equations
        #C uptake
        F1=1/CUE*(Vmax+mr)*Cmic*C/(Km+C)
        #growth respiration
        F4=(1/CUE-1)*Vmax*Cmic*C/(Km+C)
        #maintnance respiration
        F5=(1/CUE-1)*mr*Cmic*C/(Km+C)
        #microbial mortality
        F8=mr*Cmic

        #states
        dCmic<-F1-(F4+F5)-F8
        dC<--F1+F8

        return(list(c(dCmic, dC), r=F4+F5, Protc=Protc))

@@

        #equations
        #C uptake
        F1=1/CUE*(Vmax+mr)*Cmic*C/(Km+C)
        #growth respiration
        F4=(1/CUE-1)*Vmax*Cmic*C/(Km+C)
        #maintnance respiration
        F5=(1/CUE-1)*mr*Cmic*C/(Km+C)
        #microbial mortality
        F8=mr*Cmic

        #states
        dCmic<-F1-(F4+F5)-F8
        dC<--F1+F8

        return(list(c(dCmic, dC), r=F4+F5, DNAc=DNAc))

@@

@@

#results of the previous iteration (not run)
# mend_all_results_par<-rbind(mend_i1[[1]]$pars, mend_i1_DNA[[1]]$pars,
#                    mend_i2[[1]]$pars, mend_i2[[2]]$pars, mend_i2[[3]]$pars,
#                    mend_i2_DNA[[1]]$pars, mend_i2_DNA[[2]]$pars, mend_i2_DNA[[3]]$pars,
#                    mend_i3[[1]]$pars, mend_i3[[2]]$pars,
#                    mend_i3_DNA[[1]]$pars, mend_i3_DNA[[2]]$pars,
#                    mend_i4[[1]]$pars, mend_i4[[2]]$pars, mend_i4[[3]]$pars,
#                    mend_i4[[4]]$pars, mend_i4[[5]]$pars, mend_i4[[6]]$pars,
#                    mend_i4_DNA[[1]]$pars, mend_i4_DNA[[2]]$pars, mend_i4_DNA[[3]]$pars,
#                    mend_i4_DNA[[4]]$pars, mend_i4_DNA[[5]]$pars, mend_i4_DNA[[6]]$pars)
# 
# mend_all_results_good<-rbind(mend_i1$goodness[c(2,1),c(4:7)],
#                  mend_i2$goodness[c(2,1),c(4:7)],
#                  mend_i3$goodness[c(2,1),c(4:7)],
#                  mend_i4$goodness[c(2,1),c(4:7)],
#                  mend_i1_DNA$goodness[c(2,1),c(4:7)],
#                  mend_i2_DNA$goodness[c(2,1),c(4:7)],
#                  mend_i3_DNA$goodness[c(2,1),c(4:7)],
#                  mend_i4_DNA$goodness[c(2,1),c(4:7)])
# 
# mend_all_results_good[mend_all_results_good$R2<0,"R2"]<-c(0)
# mend_all_results_good$Name<-c(rep(c("Respiration rate", "Protein"), times=4),
#                               rep(c("Respiration rate", "DNA"), times=4))
# mend_all_results_good<-mend_all_results_good[,c(5,1,2,3,4)]
#  
# 
# write.csv(mend_all_results_par, file = c("../mend_all_results_par.csv"), row.names = F)
# write.csv(mend_all_results_good, file = c("../mend_all_results_good.csv"), row.names = F)

@@

>>>>

        #equations
        #C uptake
        Cu=Vmax*Cmic*C/(Km+C)
        #respiration
        r=mr*Cmic+(1-CUE)*Cu

        #Protc=Cmic*fp
        #DNAc=Cmic*fd

        #states
        dCmic<-CUE*Cu-mr*Cmic
        dC<--Cu

        return(list(c(dCmic, dC), r=r))

@@

        #select time and the measured variables 
        yhat<-select(yhat_all, c("time", "r"))
        colnames(yhat)<-c("time", "value")

        #reformat to long format data frame
        #Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        yhat$obs<-c(odeset[order(odeset$Time), c("r")])

        #add the weighting factor
        #I want to have the weighting factor to be proportional to mean of the given variable 
        yhat$weights<-rep(mean(yhat$value, na.rm = T), times=nrow(yhat))

        #match with the Yhat data frame
        #Yhat<-merge(Yhat, yweights, by.x=c("variable"), by.y=c("variable"))

        #Yhat<-Yhat[Yhat$variable!="Protc", ]

        #now, the root mean square error is calculated
        NRMSE<-with(yhat, sum((((value-obs)/mean(weights))^2), na.rm = T))

@@

        #select time and the measured variables 
        yhat<-select(yhat_all, c("time", "r"))
        colnames(yhat)<-c("time", "value")

        #reformat to long format data frame
        #Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        yhat$obs<-c(odeset[order(odeset$Time), c("r")])

        #add the weighting factor
        #I want to have the weighting factor to be proportional to mean of the given variable 
        #yhat$weights<-rep(mean(yhat$value, na.rm = T), times=nrow(yhat))

        #match with the Yhat data frame
        #Yhat<-merge(Yhat, yweights, by.x=c("variable"), by.y=c("variable"))

        #Yhat<-Yhat[Yhat$variable!="Protc", ]

        #rsquared calculation for each variable
        SSres=with(yhat, sum(((obs-value)^2), na.rm = T))
        SStot=with(yhat, sum(((obs-mean(obs, na.rm = T))^2), na.rm = T))
        ll=with(yhat, -sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        R2<-1-SSres/SStot
        N<-length(x)
        AIC<-2*N-2*ll

        Gfit<-c(R2=R2, N=N, AIC=AIC, ll=ll)

        rsq_out<-list(Yhat=yhat, Gfit=Gfit)

@@

        #rsquared calculation for each variable
        SSres=with(res$OvP, sum(((obs-value)^2), na.rm = T))
        SStot=with(res$OvP, sum(((obs-mean(obs, na.rm = T))^2), na.rm = T))
        ll=with(res$OvP, -sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        R2<-1-SSres/SStot
        N<-length(parnames)*2
        AIC<-2*N-2*ll

        res$goodness<-c(R2=R2, N=N, AIC=AIC, ll=ll)

@@

        #rsquared calculation for each variable
        SSres=with(res$OvP, sum(((obs-value)^2), na.rm = T))
        SStot=with(res$OvP, sum(((obs-mean(obs, na.rm = T))^2), na.rm = T))
        ll=with(res$OvP, -sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        R2<-1-SSres/SStot
        N<-length(parnames)*3
        AIC<-2*N-2*ll

        res$goodness<-c(R2=R2, N=N, AIC=AIC, ll=ll)

@@

        #rsquared calculation for each variable
        SSres=with(res$OvP, sum(((obs-value)^2), na.rm = T))
        SStot=with(res$OvP, sum(((obs-mean(obs, na.rm = T))^2), na.rm = T))
        ll=with(res$OvP, -sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        R2<-1-SSres/SStot
        N<-length(parnames)*6
        AIC<-2*N-2*ll

        res$goodness<-c(R2=R2, N=N, AIC=AIC, ll=ll)
        }

@@

# #results of the previous iteration (not run)
# mend_r_results_par<-rbind(mend_r1[[1]]$pars,
#                      mend_r2[[1]]$pars, mend_r2[[2]]$pars, mend_r2[[3]]$pars,
#                      mend_r3[[1]]$pars, mend_r3[[2]]$pars,
#                      mend_r4[[1]]$pars, mend_r4[[2]]$pars, mend_r4[[3]]$pars,
#                      mend_r4[[4]]$pars, mend_r4[[5]]$pars, mend_r4[[6]]$pars)
# mend_r_results_good<-as.data.frame(rbind(mend_r1$goodness,
#                     mend_r2$goodness,
#                     mend_r3$goodness,
#                     mend_r4$goodness))
# write.csv(mend_r_results_par, file = c("../mend_r_results_par.csv"), row.names = F)
# write.csv(mend_r_results_good, file = c("../mend_r_results_good.csv"), row.names = F)

@@

        #rsquared calculation for each variable
        Gfit<-Yhat %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                        SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                        ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        Gfit$R2<-with(Gfit, 1-SSres/SStot)
        Gfit$N<-length(x)
        Gfit$AIC<-with(Gfit, 2*N-2*ll)

@@

        #equations
        #C uptake
        Cu=Vmax*Cmic*C/(Km+C)
        #respiration
        r=mr*Cmic+(1-CUE)*Cu

        Protc=Cmic*fp
        #DNAc=Cmic*fd

        #states
        dCmic<-CUE*Cu-mr*Cmic
        dC<--Cu

        return(list(c(dCmic, dC), r=r, Protc=Protc))

@@

       #equations
        #C uptake
        Cu=Vmax*Cmic*C/(Km+C)
        #respiration
        r=mr*Cmic+(1-CUE)*Cu

        #Protc=Cmic*fp
        DNAc=Cmic*fd

        #states
        dCmic<-CUE*Cu-mr*Cmic
        dC<--Cu

        return(list(c(dCmic, dC), r=r, DNAc=DNAc))

@@

      #approximate parameter estimation is done by MCMC method
      par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, CUE=0.5, mr=0.01, fd=0.5, Cmic_0=0.01), 
                          lower=c(Vmax=1e-3, Km=1e-3, CUE=0, mr=1e-5, fd=0, Cmic_0=1e-5),
                          upper=c(Vmax=10, Km=100, CUE=1, mr=10, fd=1, Cmic_0=5), niter=10000)

@@

# #results of the previous iteration (not run)
# mend_all_results_par<-rbind(mend_i1[[1]]$pars, mend_i1_DNA[[1]]$pars,
#                   mend_i2[[1]]$pars, mend_i2[[2]]$pars, mend_i2[[3]]$pars,
#                   mend_i2_DNA[[1]]$pars, mend_i2_DNA[[2]]$pars, mend_i2_DNA[[3]]$pars,
#                   mend_i3[[1]]$pars, mend_i3[[2]]$pars,
#                   mend_i3_DNA[[1]]$pars, mend_i3_DNA[[2]]$pars,
#                   mend_i4[[1]]$pars, mend_i4[[2]]$pars, mend_i4[[3]]$pars,
#                   mend_i4[[4]]$pars, mend_i4[[5]]$pars, mend_i4[[6]]$pars,
#                   mend_i4_DNA[[1]]$pars, mend_i4_DNA[[2]]$pars, mend_i4_DNA[[3]]$pars,
#                   mend_i4_DNA[[4]]$pars, mend_i4_DNA[[5]]$pars, mend_i4_DNA[[6]]$pars)
# 
# mend_all_results_good<-rbind(mend_i1$goodness[c(2,1),c(4:7)],
#                 mend_i2$goodness[c(2,1),c(4:7)],
#                 mend_i3$goodness[c(2,1),c(4:7)],
#                 mend_i4$goodness[c(2,1),c(4:7)],
#                 mend_i1_DNA$goodness[c(2,1),c(4:7)],
#                 mend_i2_DNA$goodness[c(2,1),c(4:7)],
#                 mend_i3_DNA$goodness[c(2,1),c(4:7)],
#                 mend_i4_DNA$goodness[c(2,1),c(4:7)])
# 
# mend_all_results_good[mend_all_results_good$R2<0,"R2"]<-c(0)
# mend_all_results_good$Name<-c(rep(c("Respiration rate", "Protein"), times=4),
#                              rep(c("Respiration rate", "DNA"), times=4))
# mend_all_results_good<-mend_all_results_good[,c(5,1,2,3,4)]
# 
# 
# write.csv(mend_all_results_par, file = c("../mend_all_results_par.csv"), row.names = F)
# write.csv(mend_all_results_good, file = c("../mend_all_results_good.csv"), row.names = F)

@@

```{r picina example, eval=FALSE, include=TRUE}

#######################################Monod model##################################################
monod_ex<-function(data){

        Cb<-matrix(state[1:5], ncol = 1)
        C<-matrix(state[6:10], ncol = 1)

        dCb<--k*Cb+CUE*Vmax*Cb*C/(Km+C)
        dC<-k*Cb-Vmax*Cb*C/(Km+C)

        return(list(c(dCb, dC), r=(1-CUE)*Vmax*Cb*C/(Km+C), Cmic=kec*Cb))

    #define names of parameters
    parnames<-c("Vmax", "Km", "CUE", "k", "kec")

        #initial conditions
        Cmicinit=93.6/par[["kec"]]
        Cinit=315

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(Cb1=Cmicinit, Cb2=Cmicinit, Cb3=Cmicinit, Cb4=Cmicinit, Cb5=Cmicinit,
                                        C1=315, C2=315, C3=315, C4=315, C5=315), parms=par, deriv, times=c(0, 8, 16, 24, 48, 72)))

        #select time and the measured variables
        yhat<-select(yhat_all, c("time", "Cmic1", "Cmic2", "Cmic3","Cmic4", "Cmic5",
                               "C1", "C2", "C3", "C4", "C5",
                               "r1", "r2", "r3", "r4", "r5"))

        #add the measured data to a data frame
        Yhat$obs<-c(as.numeric(odeset[c(1,6,11,16, 21, 26), c("Cmic")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("Cmic")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("Cmic")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("Cmic")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("Cmic")]),
                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("DOC")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("DOC")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("DOC")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("DOC")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("DOC")]),
                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("r")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("r")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("r")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("r")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("r")]))

        #initial conditions
        Cmicinit=93.6/par[["kec"]]
        Cinit=315

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(Cb1=Cmicinit, Cb2=Cmicinit, Cb3=Cmicinit, Cb4=Cmicinit, Cb5=Cmicinit,
                                        C1=315, C2=315, C3=315, C4=315, C5=315), parms=par, deriv, times=c(0, 8, 16, 24, 48, 72)))

        #select time and the measured variables
        yhat<-select(yhat_all, c("time", "Cmic1", "Cmic2", "Cmic3","Cmic4", "Cmic5",
                               "C1", "C2", "C3", "C4", "C5",
                               "r1", "r2", "r3", "r4", "r5"))

        #add the measured data to a data frame
        Yhat$obs<-c(as.numeric(odeset[c(1,6,11,16, 21, 26), c("Cmic")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("Cmic")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("Cmic")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("Cmic")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("Cmic")]),
                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("DOC")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("DOC")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("DOC")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("DOC")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("DOC")]),
                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("r")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("r")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("r")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("r")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("r")]))

        Yhat$variable<-c(rep("Cmic", 30), rep("C", 30), rep("r", 30))

        #rsquared calculation for each variable
        Gfit<-Yhat %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
                                                        SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                        ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        Gfit$R2<-with(Gfit, 1-SSres/SStot)
        Gfit$N<-length(x)
        Gfit$AIC<-with(Gfit, 2*N-2*ll)

      #approximate parameter estimation is done by MCMC method
      par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, CUE=0.5, k=0.01, kec=0.38),
                          lower=c(Vmax=1e-5, Km=1e-2, CUE=0, k=1e-5, kec=0),
                          upper=c(Vmax=1, Km=20, CUE=1, k=10, kec=1), niter=10000)

    res<-estim(odeset=data)

#################################################MEND#######################################################
mend_ex<-function(data){

   #mend growth function
    deriv<-function(time, state, pars){

        Cb<-matrix(state[1:5], ncol = 1)
        C<-matrix(state[6:10], ncol = 1)

        dCb<-CUE*Vmax*Cb*C/(Km+C)-mr*Cb
        dC<--Vmax*Cb*C/(Km+C)

        return(list(c(dCb, dC), r=(1-CUE)*Vmax*Cb*C/(Km+C)+mr*Cb, Cmic=kec*Cb))

    #define names of parameters
    parnames<-c("Vmax", "Km", "CUE", "mr", "kec")

        #initial conditions
        Cmicinit=93.6/par[["kec"]]
        Cinit=315

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(Cb1=Cmicinit, Cb2=Cmicinit, Cb3=Cmicinit, Cb4=Cmicinit, Cb5=Cmicinit,
                                        C1=315, C2=315, C3=315, C4=315, C5=315), parms=par, deriv, times=c(0, 8, 16, 24, 48, 72)))

        #select time and the measured variables
        yhat<-select(yhat_all, c("time", "Cmic1", "Cmic2", "Cmic3","Cmic4", "Cmic5",
                               "C1", "C2", "C3", "C4", "C5",
                               "r1", "r2", "r3", "r4", "r5"))

        #add the measured data to a data frame
        Yhat$obs<-c(as.numeric(odeset[c(1,6,11,16, 21, 26), c("Cmic")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("Cmic")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("Cmic")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("Cmic")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("Cmic")]),
                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("DOC")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("DOC")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("DOC")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("DOC")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("DOC")]),
                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("r")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("r")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("r")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("r")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("r")]))

        #initial conditions
        Cmicinit=93.6/par[["kec"]]
        Cinit=315

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(Cb1=Cmicinit, Cb2=Cmicinit, Cb3=Cmicinit, Cb4=Cmicinit, Cb5=Cmicinit,
                                        C1=315, C2=315, C3=315, C4=315, C5=315), parms=par, deriv, times=c(0, 8, 16, 24, 48, 72)))

        #select time and the measured variables
        yhat<-select(yhat_all, c("time", "Cmic1", "Cmic2", "Cmic3","Cmic4", "Cmic5",
                               "C1", "C2", "C3", "C4", "C5",
                               "r1", "r2", "r3", "r4", "r5"))

        #add the measured data to a data frame
        Yhat$obs<-c(as.numeric(odeset[c(1,6,11,16, 21, 26), c("Cmic")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("Cmic")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("Cmic")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("Cmic")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("Cmic")]),
                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("DOC")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("DOC")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("DOC")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("DOC")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("DOC")]),
                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("r")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("r")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("r")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("r")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("r")]))

        Yhat$variable<-c(rep("Cmic", 30), rep("C", 30), rep("r", 30))

        #rsquared calculation for each variable
        Gfit<-Yhat %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
                                                        SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                        ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        Gfit$R2<-with(Gfit, 1-SSres/SStot)
        Gfit$N<-length(x)
        Gfit$AIC<-with(Gfit, 2*N-2*ll)

      #approximate parameter estimation is done by MCMC method
      par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, CUE=0.5, mr=0.01,kec=0.38),
                          lower=c(Vmax=1e-5, Km=1e-2, CUE=0, mr=1e-5, kec=0),
                          upper=c(Vmax=1, Km=20, CUE=1, mr=10, kec=1), niter=10000)

    res<-estim(odeset=data)

#################################################DEB#######################################################
deb_ex<-function(data){

   #deb function
    deriv<-function(time, state, pars){

      R<-matrix(state[1:5], ncol = 1)
      S<-matrix(state[6:10], ncol = 1)
      C<-matrix(state[11:15], ncol = 1)

      #C uptake
      Cu=Vmax*S*C/(Km+C)

      #udrzovani
      m=S*m0

      #mobilizace rezerv
      an=f*R-m

      #respirace
      r=m+pmax(an*(1-Yu), 0)

      #microbialni uhlik je neco z rezerv a neco ze struktur
      Cmic=fr*R+fs*S

      return(list(c(dR, dS, dC), Cmic=Cmic, r=r))

    #define names of parameters
    parnames<-c("Vmax", "Km", "m0", "f", "Yu", "fr", "fs", "Rinit")

        par<-x

        Cmicinit=93.6
        Cinit=315

        #initial R
        Rinit<-par[["Rinit"]]

        #initial S is then
        Sinit<-(Cmicinit-Rinit*par[["fr"]])/par[["fs"]]

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(R1=Rinit, R2=Rinit, R3=Rinit, R4=Rinit, R5=Rinit,
                                      S1=Sinit, S2=Sinit, S3=Sinit, S4=Sinit, S5=Sinit,
                                      C1=315, C2=315, C3=315, C4=315, C5=315), 
                                  parms=par[1:7], 
                                  deriv, 
                                  times=c(0, 8, 16, 24, 48, 72)))
         #select time and the measured variables 
         yhat<-select(yhat_all, c("time", "Cmic1", "Cmic2", "Cmic3","Cmic4", "Cmic5",
                               "C1", "C2", "C3", "C4", "C5",
                               "r1", "r2", "r3", "r4", "r5"))

        #add the measured data to a data frame
        Yhat$obs<-c(as.numeric(odeset[c(1,6,11,16, 21, 26), c("Cmic")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("Cmic")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("Cmic")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("Cmic")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("Cmic")]),
                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("DOC")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("DOC")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("DOC")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("DOC")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("DOC")]),
                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("r")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("r")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("r")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("r")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("r")]))

        par<-x

        #initial conditions
        Cmicinit=93.6
        Cinit=315

        #initial R
        Rinit<-par[["Rinit"]]

        #initial S is then
        Sinit<-(Cmicinit-Rinit*par[["fr"]])/par[["fs"]]

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(R1=Rinit, R2=Rinit, R3=Rinit, R4=Rinit, R5=Rinit,
                                      S1=Sinit, S2=Sinit, S3=Sinit, S4=Sinit, S5=Sinit,
                                      C1=315, C2=315, C3=315, C4=315, C5=315), 
                                  parms=par[1:7], 
                                  deriv, 
                                  times=c(0, 8, 16, 24, 48, 72)))
         #select time and the measured variables 
         yhat<-select(yhat_all, c("time", "Cmic1", "Cmic2", "Cmic3","Cmic4", "Cmic5",
                               "C1", "C2", "C3", "C4", "C5",
                               "r1", "r2", "r3", "r4", "r5"))

        #add the measured data to a data frame
        Yhat$obs<-c(as.numeric(odeset[c(1,6,11,16, 21, 26), c("Cmic")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("Cmic")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("Cmic")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("Cmic")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("Cmic")]),
                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("DOC")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("DOC")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("DOC")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("DOC")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("DOC")]),
                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("r")]),
                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("r")]),
                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("r")]),
                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("r")]),
                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("r")]))

        Yhat$variable<-c(rep("Cmic", 30), rep("C", 30), rep("r", 30))

        #rsquared calculation for each variable
        Gfit<-Yhat %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
                                                        SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                        ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        Gfit$R2<-with(Gfit, 1-SSres/SStot)
        Gfit$N<-length(x)
        Gfit$AIC<-with(Gfit, 2*N-2*ll)

      #approximate parameter estimation is done by MCMC method
      par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, m0=1e-3, f=1e-2, Yu=0.5, fr=0.8, fs=0.38, Rinit=40),
                          lower=c(Vmax=1e-5, Km=1e-2, m0=1e-5, f=1e-5, Yu=0, fr=0, fs=0, Rinit=93*0.01),
                          upper=c(Vmax=1, Km=20, m0=10, f=10, Yu=1, fr=1, fs=1, Rinit=93*0.9), niter=10000)

    res<-estim(odeset=data)
